### 2025-05-04 doc/declaration.tex
Declaration template - submit

\newpage
\thispagestyle{empty}  % <--- Hides page number

\section*{Declaration of Originality}

I hereby declare that the present seminar paper and the work reported herein was composed by and originated entirely from me without any help. All sources used from published or unpublished work of others are reported in the list of references. All parts of my work that are based on others’ work are cited as such.

This seminar paper has not been submitted for any degree or other purposes, neither at the Humboldt-Universität zu Berlin nor at any other university or college.

\vspace{20mm}

Berlin, \hrulefill

\vspace{30mm}

\hrulefill \\
Name and surname \hspace{20mm} Signature