### 2025-05-04 doc/declaration.tex
Declaration - signature

\newpage
\thispagestyle{empty}  % <--- Hides page number

\section*{Declaration of Academic Honesty}

I, Max Mustermann, hereby declare that I have not previously submitted the present work for other examinations. I wrote this work independently. All sources, including sources from the Internet, that I have reproduced in either an unaltered or modified form (particularly sources for texts, graphs, tables and images), have been acknowledged by me as such.

I understand that violations of these principles will result in proceedings regarding deception or attempted deception.

\vspace{2cm}

\rule{60mm}{0.4pt}

\vspace{0.5cm}

\text{Max Mustermann}

Berlin, \today