### 2025-05-04 doc/acknowledgements.tex

\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}  % Or adjust based on your layout
\section*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

I would like to thank my professors, colleagues, and friends for their support and insights throughout the preparation of this thesis. Special thanks go to [Name], whose guidance was invaluable during the project.

\clearpage