### 2025-05-04 doc/titlepage.tex
Title page done

\begin{titlepage}
    \begin{center}
    \vspace*{0.5cm}
    \LARGE{How Much New Information Is There in Earnings?}\\[3cm]
    \Large{Corporate Decision-Making and Quantitative Analysis}\\[1cm]
    \normalsize{
    Marketing Seminar\\
    Summer Term 2025 \\
    School of Business and Economics\\
    Humboldt University Berlin \\[3cm]
    }

    Submitted by:\\
    Max Mustermann (Matriculation no.: 123456)\\[3cm]

    Submitted to:\\
    Prof. Dr. Anastasia Danilov\\[2cm]

    Berlin, \today
    \end{center}
    \end{titlepage}