### 2025-05-05 doc/abstract.tex
Adjust abstract&titlepage to uni guidelines

\clearpage
\pagenumbering{Roman}
\setcounter{page}{3}  % Optional: set to next page after acknowledgements
\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}

This thesis examines how much new information is conveyed in quarterly earnings announcements, replicating and extending the study by Ball and Shivakumar (2008). Using a reproducible workflow based on the TRR 266 Template, I analyze U.S. and Canadian data from 1972 to 2023, applying an event-study design to assess the explanatory power of earnings windows on annual stock returns. The findings show that earnings informativeness varies across time and jurisdictions, with evidence suggesting institutional factors influence market reactions. A non-archival survey design is also proposed to explore investor perceptions, enriching the archival analysis with complementary behavioral insights.

\clearpage