### 2025-05-05 doc/titlepage.tex
Structure Aligned with Accounting Chair Requieremnts

\begin{titlepage}
    \thispagestyle{empty}

    \vspace*{2cm}

    % Thesis title (2-line format)
    {\Large
    \noindent How Much New Information Is There in Earnings?\\
    Evidence from Earnings Releases in the U.S. and Canada}

    \vspace{2cm}

    % Type of thesis - largest font (not bold)
    {\LARGE
    \noindent
    Master’s thesis}

    \vspace{2cm}

    % Degree info
    {\normalsize
    \noindent for acquiring the degree of\\
    Master of Science (M.Sc.)

    \vspace{0.5cm}

    \noindent in Business Administration

    \vspace{0.5cm}

    \noindent at the School of Business and Economics\\
    of Humboldt-Universität zu Berlin

    \vspace{2cm}

    % Submission info
    \noindent submitted by\\
    Max Mustermann\\
    Student no. 1234567

    \vspace{2cm}

    \noindent First Examiner: << name of first examiner >>\\
    Second Examiner: << name of second examiner >>

    \vspace{1cm}

    \noindent Berlin, \today
    }

\end{titlepage}