### 2025-05-23 doc/paper.qmd
paper done; analysis adjusted

---
format:
  pdf:
    documentclass: article
    number-sections: true
    toc: true
    lof: true
    lot: true
    include-before-body:
      - titlepage.tex
      - preamble.tex  # This sets Roman page numbering before the TOC
    include-after-body:
      - declaration.tex

bibliography: references.bib
biblio-style: apsr

fig_caption: yes
fontsize: 12pt # 12-point font acc. to Accounting guidelines
ident: yes
always_allow_html: yes
number-sections: true
header-includes:
  - \usepackage[nolists]{endfloat} # if commenting out, you place figures&tables inside text
  - \usepackage{float} # in order to place tables correctly
  - \floatplacement{table}{H} # in order to place tables correctly
  - \usepackage{setspace}\onehalfspacing # 1.5 line spacing acc. to Accounting guidelines
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \usepackage[hang,flushmargin]{footmisc}
  - \usepackage{caption}
  - \captionsetup[table]{skip=15pt,font=bf} # was 24pt initially
  - \captionsetup[figure]{skip=15pt, font=bf} # consistency with tables
  - \usepackage{array}
  - \usepackage{threeparttable}
  - \usepackage{adjustbox}
  - \usepackage{graphicx}
  - \usepackage{csquotes}
  - \usepackage{indentfirst}
  - \usepackage[margin=1in]{geometry}
  - \usepackage{ragged2e} #for declaration intend
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}  % clear headers/footers
  - \renewcommand{\headrulewidth}{0pt}
  - \renewcommand{\footrulewidth}{0pt}
  - \fancyfoot[R]{\makebox[0pt][r]{\thepage\hspace{1in}}} % right-aligned footer # page numbering on lower right margin acc. to WiWi guidelines
  - \setlength{\footskip}{0cm} # acc. to Wiwi guidelines - control a 1 to 1.5 cm gap below page numbers
---

\pagebreak

```{python}
#| echo: false
#| output: false

import pickle
import pandas as pd

# Load the prepared summary table
with open('../output/summary_statistics.pickle', 'rb') as f:
    summary = pickle.load(f)

# Clean column names for LaTeX rendering
def escape_special_characters(df):
    df.columns = [col.replace('_', '\\_').replace('#', '\\#') for col in df.columns]
    return df

# Formatting for float values
def custom_float_format(x):
    return f"{x:.2f}" if pd.notnull(x) else ""

# Function to prep LaTeX table
# Function to prep LaTeX table with optional description
def prep_latex_table(df, caption=None, label=None, description=None):
    df = escape_special_characters(df.reset_index(drop=True))
    num_columns = len(df.columns)
    column_format = 'l' + 'r' * (num_columns - 1)

    latex_table = df.to_latex(
        column_format=column_format,
        index=False,
        float_format=custom_float_format,
        escape=False
    )

    latex_caption = f"\\caption{{{caption}}}" if caption else ""
    latex_description = f"\\caption*{{\\textnormal{{{description}}}}}" if description else ""
    latex_label = f"\\label{{{label}}}" if label else ""

    latex_table_lines = [
        "\\begin{table}",
        "\\centering",
        "\\fontsize{12pt}{16pt}\\selectfont", #  12pt to adjust font size acc. to guidelines, 16pt for 1.5 line spacing in table
        "\\resizebox{\\textwidth}{!}{%",
        latex_table,
        "}",  # End resizebox
        latex_caption,
        latex_description,
        latex_label,
        "\\end{table}"
    ]
    return "\n".join(line for line in latex_table_lines if line)
```

\clearpage
# Acknowledgements {-}
I would like to thank my professors, colleagues, and friends for their support and insights throughout the preparation of this thesis. Special thanks go to [Name], whose guidance was invaluable during the project.

\clearpage

# Abstract {-}
This thesis template explores global development patterns using the publicly available Gapminder dataset, with a specific focus on the year 2007. Leveraging a reproducible workflow adapted from the TRR 266 Template, the project presents descriptive statistics by continent and a visual analysis of the relationship between life expectancy, income per capita, and population. The findings reveal clear disparities in health and wealth outcomes across world regions. A comparison with Gapminder’s 2021 global development chart highlights persistent global inequalities, while illustrating notable progress in certain regions. The analysis serves as a modern, transparent example of public data storytelling in empirical research.

\newpage

# List of Abbreviations {-}
\setlength{\parindent}{0em}
\begin{tabular}{@{}ll}
\textbf{GDP}   & Gross Domestic Product \\
\textbf{TREAT} & TRR 266 Template for Reproducible Empirical Accounting Research \\
\end{tabular}

\newpage
\pagenumbering{arabic}
\setlength{\parindent}{4em}
\setlength{\parskip}{0em}
\newgeometry{left=6cm, right=1.5cm, top=2cm, bottom=2cm}

# Introduction {#sec-introduction}
This project serves as a template-based example for conducting and documenting reproducible empirical data analysis using publicly available datasets. It leverages the TRR 266 Template for Reproducible Empirical Accounting Research (TREAT) and builds upon the methodological foundations of the Corporate Decision-Making and Quantitative Analysis course, as well as the Accounting Reading Group. To support future empirical thesis projects, this template was developed in adherence to the formal formatting and content guidelines of the Institute of Accounting and Auditing at the School of Business and Economics, Humboldt-Universität zu Berlin. It is not a completed thesis but a demonstrative project hosted in the [template repository](https://github.com/melmzv/huwiwiQuartoThesis).

This template project explores global development patterns using the Gapminder dataset, focusing on the year 2007. The project illustrates how reproducible research workflows can be applied by using public and open-source data. Specifically, it presents a summary statistics table by continent and visualizes the relationship between GDP per capita and life expectancy, offering a static snapshot of world development in 2007. These descriptive insights are presented in @sec-task1.

The results are further contrasted with Gapminder’s interactive global development visualizations as of 2021 [@gapminder_health_chart]. This comparative discussion is provided in @sec-task2. The project concludes with reflections on transparency, reproducibility, and the importance of making research accessible to a broader audience, as discussed in @sec-conclusion.

# Descriptive Analysis of Global Development in 2007 {#sec-task1}
This section explores global development patterns using Gapminder data for the year 2007. It presents a continent-level summary statistics table and a scatter plot showing the relationship between GDP per capita and life expectancy. These outputs serve as a static snapshot of global well-being during that year and demonstrate how transparent and reproducible workflows can be applied to public datasets.

## Research Design Choices and Assumptions {#sec-research_design_assumptions1}
In line with @gapminder_health_chart, the analysis uses the Gapminder dataset available via the `gapminder` Python package. This dataset contains country-level information on population, GDP per capita (inflation-adjusted), and life expectancy for various years in five-year intervals. For this project, I focus on the year 2007 - the most recent year available in the dataset at the time.

To ensure clarity and consistency, several assumptions are applied. The analysis is restricted to the year 2007, assuming that cross-sectional variation in life expectancy and GDP per capita during that year sufficiently captures key development patterns. No imputation is performed; only complete observations are included. GDP per capita is used in inflation-adjusted international dollars, as provided by @gapminder_health_chart, without applying additional currency normalization. These choices prioritize interpretability and reproducibility, while acknowledging that dynamic trends and structural differences across regions are not captured in this static snapshot. All data originates from the Gapminder Foundation, which curates and harmonizes publicly available indicators to promote global development awareness [@gapminder_health_chart].

These assumptions, together with the procedural details in @sec-replication_steps1, guide the replication and ensure transparency in design and implementation.

## Replication Steps {#sec-replication_steps1}
This section outlines the modular workflow used to generate the summary table and figure, aligned with @gapminder_health_chart.

**Step 1: Pulling the Data**  
The data is loaded from the `gapminder` Python package. The 2007 subset is saved in `.parquet` format for modern data handling and versioning.

**Step 2: Data Preparation**  
The dataset is checked for missing values, which are reported and excluded. All numerical columns are rounded to two decimal places. No transformations, sorting, or new variables are added to preserve the original data structure.

**Step 3: Analysis Implementation and Reproduction of Tables and Figure**  
In the final step, two key outputs are generated to illustrate global development patterns in 2007. First, a summary table is produced, aggregating life expectancy, GDP per capita, and total population figures by continent. This table provides a concise overview of regional development differences. Second, a scatter plot is created to visualize the relationship between GDP per capita and life expectancy, where each country’s bubble size corresponds to its population and color represents its continent.

## Results {#sec-results1}
This section presents a static snapshot of global development in 2007 using the Gapminder dataset. The figure and table illustrate clear regional disparities in wealth and health outcomes - highlighting, for example, the relatively high life expectancy in Europe and Oceania compared to lower-income regions like Africa.

@tbl-gapminder2007_summary provides aggregated statistics by continent and complements the visual by quantifying central tendencies and demographic scale. It presents average and median values for life expectancy and GDP per capita, along with total population and number of countries represented per continent in 2007. A total of `{python} f"{summary['Num_Countries'].sum():,}"` countries are included in the 2007 snapshot.

```{python}
#| label: tbl-gapminder2007_summary
#| tbl-cap: "Summary Statistics by Continent (Gapminder 2007)"
#| echo: false
#| output: asis

print(prep_latex_table(
    summary,
    description="\\fontsize{10pt}{12pt}\\selectfont This table reports the mean and median life expectancy and GDP per capita, along with the total population and number of countries per continent for the year 2007."
))
```

@fig-gapminder2007_plot illustrates a scatter plot based on Gapminder’s 2007 data, mapping GDP per capita against life expectancy. Each country is represented as a bubble whose size corresponds to its population and whose color indicates its continent. The plot reveals a positive relationship between economic prosperity and life expectancy.

```{python}
#| label: fig-gapminder2007_plot
#| fig-cap: "Gapminder 2007: Life Expectancy vs. GDP per Capita"
#| echo: false

from IPython.display import Image
Image(filename="../output/gapminder_figure.png")
```

Together, these outputs provide a clear and interpretable overview of global development disparities, offering insights into how income and health outcomes align across continents at a given point in time.

# Global Development Patterns in 2021 {#sec-task2}
This section compares the 2007 snapshot with Gapminder’s 2021 visualization to assess whether key patterns - like the link between GDP per capita and life expectancy - persist and to highlight shifts in global inequality and regional development.

## Comparative Assumptions and Design Choices {#sec-research_design_assumptions3}
This section outlines the assumptions for comparing the 2007 Gapminder dataset to a visualization of global development patterns in 2021.

The assumptions guiding the comparison are as follows: visual inspection is used to assess trends in life expectancy and income per capita; bubble size reflects each country’s population, as defined by Gapminder; and no new calculations are applied - the 2021 snapshot displays curated values from the Gapminder Foundation [-@gapminder_health_chart].

The relationship between country-level variables and their graphical representation in the visualizations can be summarized as shown in @eq-bubble_mapping:
$$
\text{Bubble}_{i} = f\left(\text{GDP}_i,\ \text{LifeExp}_i,\ \text{Pop}_i,\ \text{Continent}_i\right)
$$ {#eq-bubble_mapping}

## Results {#sec-results3}
This section compares the 2007 snapshot, rendered from Gapminder data in Python in @sec-task1, with Gapminder’s official visualization of global development in 2021 presented below.
```{python}
#| label: fig-gapminder_original
#| fig-cap: "Original World Health Chart 2021 from Gapminder Foundation"
#| echo: false

from IPython.display import Image
Image(filename="countries_health_wealth_2021-1.png")
```

# Conclusion {#sec-conclusion}
This thesis template demonstrates how the TRR 266 framework can be applied to structure a reproducible and transparent empirical accounting study. Using Quarto proved to be a particularly effective and user-friendly solution for structuring and rendering the thesis. It facilitates seamless integration of code, results, and narrative, making it an ideal tool for students at the Institute of Accounting and Auditing at the School of Business and Economics, HU Berlin. Thanks for reading!

\pagebreak

\restoregeometry
\setcounter{table}{0}
\renewcommand{\thetable}{\arabic{table}}

# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent